Syntax and Semantics of Abstract Binding Trees

### 2015-11-09 paper.tex
cite pfpl
Simplify Grothendieck construction explanation

\newcommand\ArrId[1]{1_{#1}}
\newcommand\Match[2]{\IMode{#1}\equiv\OMode{#2}}

@@

\newcommand\Cats{\mathbf{Cat}}
\newcommand\Discrete[1]{#1_{\equiv}}
\newcommand\Arr[1]{\mathbf{{#1}^{\rightarrow}}}
\newcommand\Sets{\mathbf{Set}}
\newcommand\SCtx{\mathbf{SCtx}}
\newcommand\Ctx{\mathbf{Ctx}}
\newcommand\Seq{\mathbf{Seq}}
\newcommand\HypCat{\mathbf{H}}

@@

Then, fix a covariant presheaf (copresheaf) of operators
$\Of{\Operators}{\SCtx\times\Discrete{\Arities}\to\Sets}$ such that the arrows
in $\SCtx$ lift to renamings of operators' parameters. Via the Grothendieck
construction\footnote{In this case, $\mathbf{C}\Grothendieck{}{\Psi}$ represents
  the category of elements of a copresheaf $\Of{\Psi}{\mathbf{C}\to\Sets}$ but
  we keep the $\mathbf{C}$ implicit and simply refer to it as the Grothendieck
  construction. Alternatively, this construction can be understood as a coend
  $\IMode{\mathbf{C}\Grothendieck{}{\Psi}}\cong \IMode{\int^{c \in
      \mathbf{C}}\Coslice{c}{\mathbf{C}} \otimes \Discrete{\Psi(c)}}$.}
$\Of{\Grothendieck{}{(-)}}{\Sets^{\mathbf{C}}\to\Cats}$ on operators, we have a
category of objects
$\Member{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Grothendieck{}{\Operators}}$ for
$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and morphisms
$\Hom{\Grothendieck{}{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Pair{\Pair{\Upsilon'}{a'}}{\vartheta'}}$
for $\Of{\Pair{\varrho}{f}}{\Pair{\Upsilon}{a}\to\Pair{\Upsilon'}{a'}}$ such
that
$\EqMember{\FLift{\varrho}{f}(\vartheta)}{\vartheta'}{\Operators\Pair{\Upsilon'}{a'}}$.
Equivalently, $\Grothendieck{}{\Operators}$ is the pullback of $\Operators$
along the codomain Grothendieck bifibration $\pi_{cod}$ where
$\pi_{\Operators}$ is also a discrete Grothendieck opfibration:

\[
\begin{tikzcd}[ampersand replacement = \&, sep = large]
  {\OMode{\Grothendieck{}{\Operators}}}
  \arrow[rr,
    "{\IMode{\vartheta\in\Operators\Pair{\Upsilon}{a}\ \mapsto\
      \ArrId{\Operators\Pair{\Upsilon}{a}}}}"] \arrow[d, "\IMode{\pi_{\Operators}}" left]
  \& \& {\IMode{\Arr{\Sets}}}
    \arrow[d, "\IMode{\pi_{cod}}"] \\
  {\IMode{\SCtx\times\Discrete{\Arities}}} \arrow[rr, "\IMode{\Operators}" below] \&
  \& {\IMode{\Sets}} \\
\end{tikzcd}
\]

@@

Having specified an approximation of the grammar of telescopes and sequents in
the abt logical framework, we can proceed to define proper wellformedness via
\emph{inductive refinement}. The basic idea is to introduce a new form of
(meta)-judgment $\IsWf{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ which expresses the
extrinsic wellformedness properties we wish to verify, presupposing
$\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we introduce an
analogous judgment on bound terms, $\IsWf{\Omega}{\Upsilon}{\Gamma}{E}{v}$
presupposing $\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as
follows:
\[
  \infer{
    \IsWf{\Omega}{\Upsilon}{\Gamma}{
      \MkBTm{\vec{u}}{\vec{x}}{M}
    }{
      \MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}
    }
  }{
    \IsWf{\Omega}{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma,\vec{x}:\vec{\tau}}{M}{\tau}
  }
\]

>>>>

\newcommand\ArrId[1]{1_{#1}}
\newcommand\Shriek[1]{\,{{{\mathexclam}}}_{#1}}
\newcommand\Match[2]{\IMode{#1}\equiv\OMode{#2}}

@@

\newcommand\Iso[2]{#1\mathrel{{{\cong}}}#2}
\newcommand\Cats{\mathbf{Cat}}
\newcommand\Discrete[1]{#1_{\equiv}}
\newcommand\Arr[1]{\mathbf{{#1}^{\rightarrow}}}
\newcommand\Sets{\mathbf{Set}}
\newcommand\SCtx{\mathbf{SCtx}}
\newcommand\Ctx{\mathbf{Ctx}}
\newcommand\Seq{\mathbf{Seq}}
\newcommand\HypCat{\mathbf{H}}

@@

Then, fix a covariant presheaf (copresheaf) of operators
$\Of{\Operators}{\SCtx\times\Discrete{\Arities}\to\Sets}$ such that the arrows
in $\SCtx$ lift to renamings of operators' parameters. Then by the Grothendieck
construction $\Iso{\Grothendieck{}{\Operators}}{\Comma{𝟙}{\Operators}}$
we have a category of elements of $\Operators$ where objects are triples
$\Pair{\Pair{\Upsilon}{a}}{\vartheta}$, interpreted as
$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$, and morphisms are diagrams:

\[
  \begin{tikzcd}[ampersand replacement = \&]
    \& 𝟙
      \arrow[dl, "\vartheta", swap]
      \arrow[dr, "\vartheta'"]
    \&
    \\
    {\Operators\Pair{\Upsilon}{a}}
      \arrow[rr, "\Operators\Pair{\rho}{f}", swap]
    \&
    \& {\Operators\Pair{\Upsilon'}{a'}}
    \\
    {\Pair{\Upsilon}{a}}
      \arrow[rr, "\Pair{\rho}{f}", swap]
    \&
    \& {\Pair{\Upsilon'}{a'}}
  \end{tikzcd}
\]\\

@@

Having specified an approximation of the grammar of telescopes and sequents in
the abt logical framework, we can proceed to define proper wellformedness via
\emph{inductive refinement} \cite{harper:2016}. The basic idea is to introduce a new form of
(meta)-judgment $\IsWf{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ which expresses the
extrinsic wellformedness properties we wish to verify, presupposing
$\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we introduce an
analogous judgment on bound terms, $\IsWf{\Omega}{\Upsilon}{\Gamma}{E}{v}$
presupposing $\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as
follows:
\[
  \infer{
    \IsWf{\Omega}{\Upsilon}{\Gamma}{
      \MkBTm{\vec{u}}{\vec{x}}{M}
    }{
      \MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}
    }
  }{
    \IsWf{\Omega}{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma,\vec{x}:\vec{\tau}}{M}{\tau}
  }
\]
